---
title: "Randomly select students for lectures with limited seats"
output: 
  flexdashboard::flex_dashboard:
    css: "style.css"
    source_code: "https://github.com/ICTO-FMG/shiny-app-limited-seats"
    social: [ "twitter", "facebook", "linkedin" ]
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shinyjs)
```

Column {.sidebar}
-----------------------------------------------------------------------

#### Inputs:

```{r}
useShinyjs(rmd = TRUE)
textAreaInput(inputId = "student_data", "Paste studentnumbers below each other", height = "100px", placeholder = "1234\n2345\n3456")
numericInput(inputId = 'N_lectures', 'Choose the number of lectures', 1, min = 1)
numericInput(inputId = 'N_seats', 'Choose the number of available seats per lecture', 1, min = 1)
actionButton(inputId = 'submit_button', 'GO')
```

#### Privacy

This tool does not store any privacy information on the user nor the data that is processed.

Column {data-width=700 .tabset}
-----------------------------------------------------------------------

### Instruction

Add all student or use the Microsoft forms template below to have students that want to attend, register for on campus lectures prior to the start of your course. Just hit the duplicate button on the top right to use it for yourself, and use the template text from the tab obove.

* <a href="https://forms.office.com/Pages/ShareFormPage.aspx?id=zcrxoIxhA0S5RXb7PWh05WYE6t316eFCnfdGp1GENPBURUowT1U4MFczOEJZWDhINVpKWFBJNlczMC4u&sharetoken=EG3pTHmtlWdskAT36OYf" class="special" target="_blank">MS forms template</a>

When you have collected all student numbers for those who want to attend on campus lectures, just paste all student numbers into the box on the left, indicate the number of lectures for your course and the number of available seats per lecture, and hit GO.

A download button will apear to the right. Just download the excel file and use the student numbers per lecture to inform your students using BCC in [outlook on the web](http://outlook.office.com).

### Template text

#### Template

Dear Students,

Due to the COVID restrictions, there are only a limited number of seats available for the lectures. Could you indicate if you are interested in attending the live lectures. If so, please let me know through the form below before [add date here].

* [add link to form here] 

We will randomly assign students to the available slots and inform those selected through email.

Column {data-width=300}
-----------------------------------------------------------------------

### N lectures {.value-box}

```{r, echo=FALSE}
renderValueBox({
valueBox(as.numeric(input$N_lectures), icon = "fa-chalkboard-teacher", caption = "Lectures" )
})
```

### N Seats {.value-box}

```{r, echo=FALSE}
renderValueBox({
valueBox(as.numeric(input$N_seats), 
         icon    = "fa-couch", 
         caption = "Seats", 
         color   = "orange")
})
```

### Download

When ready, an excel file can be downloaded by clicking on the button below.

```{r}
shinyjs::hidden(downloadLink('download_table', 'Download table', class="btn btn-default download"))
tableOutput(outputId = 'result')
students <- eventReactive(input$submit_button, {(input$student_data)})
  
  results <- eventReactive(input$submit_button, {
    
    N_spots <- input$N_lectures*input$N_seats
    students_vector <- strsplit(students(), "\n")[[1]]
    N_students <- length(students_vector)
    
    if(N_students > N_spots) {
      
      sampled_students <- sample(students_vector, size = N_spots, replace = F)
      df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
      
    } else if (N_students < input$N_seats) { 
      
      df_results <- as.data.frame(matrix(students_vector, ncol = input$N_lectures, nrow = N_students))
      
    } else {
      
      N_visits_every_stud <- N_spots %/% N_students # number of times every student may come
      remaining_spots <- N_spots %% N_students # remaining number of spots to sample for
      
      remaining_spots_sample <- sample(students_vector, size = remaining_spots, replace = F)
      sampled_students <- c(rep(students_vector, times= N_visits_every_stud), remaining_spots_sample)
      
      while(anyDuplicated(tail(sampled_students, input$N_seats)) != 0) {
        
        remaining_spots_sample <- sample(students_vector, size = remaining_spots, replace = F)
        sampled_students <- c(rep(students_vector, times= N_visits_every_stud), remaining_spots_sample)
        
      } # if last lecture contains duplicates, take a new sample
      
      df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
      
    }
    
    colnames(df_results) <- paste0("Lecture ", 1:input$N_lectures)
    df_results
    
  })
  
 # output$result <- renderTable({results()})
  
  output$download_table <- downloadHandler(
    filename = function(){paste("lecture-seats-sampler", Sys.Date(), ".xlsx", sep="")},
    content = function(file){writexl::write_xlsx(results(), file)},
    contentType = "text/csv")
  
  observeEvent(input$submit_button, {
    if (input$submit_button == FALSE){
      shinyjs::hide("download_table")}
    else {shinyjs::show("download_table")}
  })
```
