---
title: "Randomly select students for lectures with limited seats"
output: 
  flexdashboard::flex_dashboard:
    css: "style.css"
    source_code: "https://github.com/ICTO-FMG/shiny-app-limited-seats"
    social: [ "twitter", "facebook", "linkedin" ]
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shinyjs)
```

Column {.sidebar}
-----------------------------------------------------------------------

#### Inputs:

```{r}
useShinyjs(rmd = TRUE)
textAreaInput(inputId = "student_data", "Paste studentnumbers below each other", height = "100px", placeholder = "1234\n2345\n3456")
numericInput(inputId = 'N_lectures', 'Choose the number of lectures', 1, min = 1)
numericInput(inputId = 'N_seats', 'Choose the number of available seats per lecture', 1, min = 1)
actionButton(inputId = 'submit_button', 'GO')
```

Column {data-width=700}
-----------------------------------------------------------------------

### Instruction

Use the Microsoft forms template below to have students register for on campus lectures prior to the start of your course. Just hit the duplicate button on the top right to use it for yourself.

> [MS forms template]( https://forms.office.com/Pages/ShareFormPage.aspx?id=zcrxoIxhA0S5RXb7PWh05WYE6t316eFCnfdGp1GENPBURUowT1U4MFczOEJZWDhINVpKWFBJNlczMC4u&sharetoken=EG3pTHmtlWdskAT36OYf)

When you have collected all student numbers for those who want to attend on campus lectures, just paste all student numbers into the box on the right, indicate the number of lectures for your cours and the number of available seats per lecture, and hit go.

A download button will apear here to the right. Just download the excel file and use the student numbers per lecture to inform your students using BCC in outlook.

Column {data-width=300}
-----------------------------------------------------------------------

### N lectures {.value-box}

```{r, echo=FALSE}
renderValueBox({
valueBox(as.numeric(input$N_lectures), icon = "fa-school", caption = "N Lectures" )
})
```

### N Seats {.value-box}

```{r, echo=FALSE}
renderValueBox({
valueBox(as.numeric(input$N_seats), icon = "fa-chair", caption = "N Seats", color = "orange")
})
```

### Download

When ready, an excel file can be downloaded by clicking on the button below.

```{r}
shinyjs::hidden(downloadLink('download_table', 'Download table'))
tableOutput(outputId = 'result')
students <- eventReactive(input$submit_button, {(input$student_data)})
  
  results <- eventReactive(input$submit_button, {
    
    N_spots <- input$N_lectures*input$N_seats
    students_vector <- strsplit(students(), "\n")[[1]]
    N_students <- length(students_vector)
    
    if(N_students > N_spots) {
      
      sampled_students <- sample(students_vector, size = N_spots, replace = F)
      df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
      
    } else if (N_students < input$N_seats) { 
      
      df_results <- as.data.frame(matrix(students_vector, ncol = input$N_lectures, nrow = N_students))
      
    } else {
      
      N_visits_every_stud <- N_spots %/% N_students # number of times every student may come
      remaining_spots <- N_spots %% N_students # remaining number of spots to sample for
      
      remaining_spots_sample <- sample(students_vector, size = remaining_spots, replace = F)
      sampled_students <- c(rep(students_vector, times= N_visits_every_stud), remaining_spots_sample)
      
      while(anyDuplicated(tail(sampled_students, input$N_seats)) != 0) {
        
        remaining_spots_sample <- sample(students_vector, size = remaining_spots, replace = F)
        sampled_students <- c(rep(students_vector, times= N_visits_every_stud), remaining_spots_sample)
        
      } # if last lecture contains duplicates, take a new sample
      
      df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
      
    }
    
    colnames(df_results) <- paste0("Lecture ", 1:input$N_lectures)
    df_results
    
  })
  
 # output$result <- renderTable({results()})
  
  output$download_table <- downloadHandler(
    filename = function(){paste("lecture-seats-sampler", Sys.Date(), ".xlsx", sep="")},
    content = function(file){writexl::write_xlsx(results(), file)},
    contentType = "text/csv")
  
  observeEvent(input$submit_button, {
    if (input$submit_button == FALSE){
      shinyjs::hide("download_table")}
    else {shinyjs::show("download_table")}
  })
```
