---
title: "Randomly select students for lectures with limited seats"
output: 
  flexdashboard::flex_dashboard:
    css: "style.css"
    source_code: "https://github.com/ICTO-FMG/shiny-app-limited-seats"
    social: [ "twitter", "facebook", "linkedin" ]
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shinyjs)
```

Column {.sidebar}
-----------------------------------------------------------------------

#### Inputs:

```{r}
useShinyjs(rmd = TRUE)
textAreaInput(inputId = "student_data", "Paste studentnumbers below each other", height = "100px", placeholder = "1234\n2345\n3456")
numericInput(inputId = 'N_lectures', 'Choose the number of lectures', 1, min = 1)
numericInput(inputId = 'N_seats', 'Choose the number of available seats per lecture', 1, min = 1)
actionButton(inputId = 'submit_button', 'GO')
```

#### Privacy

This tool does not store any privacy information on the user nor the data that is processed.

Column {data-width=700 .tabset}
-----------------------------------------------------------------------

### Instruction

Add all student or use the Microsoft forms template below to have students that want to attend, register for on campus lectures prior to the start of your course. Just hit the duplicate button on the top right to use it for yourself, and use the template text from the tab obove.

* <a href="https://forms.office.com/Pages/ShareFormPage.aspx?id=zcrxoIxhA0S5RXb7PWh05WYE6t316eFCnfdGp1GENPBURUowT1U4MFczOEJZWDhINVpKWFBJNlczMC4u&sharetoken=EG3pTHmtlWdskAT36OYf" class="special" target="_blank">MS forms template</a>

When you have collected all student numbers for those who want to attend on campus lectures, just paste all student numbers into the box on the left, indicate the number of lectures for your course and the number of available seats per lecture, and hit GO.

A download button will apear to the right. Just download the excel file and post it on canvas or use the student numbers per lecture to inform your students using BCC in [outlook on the web](http://outlook.office.com).

### Template text

#### Template

Dear Students,

Due to the COVID restrictions, there are only a limited number of seats available for the lectures. Could you indicate if you are interested in attending the live lectures. If so, please let me know through the form below before [add date here].

* [add link to form here] 

We will randomly assign students to the available slots and inform those selected through email.

Column {data-width=300}
-----------------------------------------------------------------------

### N lectures {.value-box}

```{r, echo=FALSE}
renderValueBox({
valueBox(as.numeric(input$N_lectures), icon = "fa-chalkboard-teacher", caption = "Lectures" )
})
```

### N Seats {.value-box}

```{r, echo=FALSE}
renderValueBox({
valueBox(as.numeric(input$N_seats), 
         icon    = "fa-couch", 
         caption = "Seats", 
         color   = "orange")
})
```

### Download

When ready, an excel file can be downloaded by clicking on the button below.

```{r}
shinyjs::hidden(downloadLink('download_table', 'Download table', class="btn btn-default download"))
tableOutput(outputId = 'result')
students <- eventReactive(input$submit_button, {(input$student_data)})
  
  results <- eventReactive(input$submit_button, {
    
    N_spots <- input$N_lectures*input$N_seats
    students_vector <- strsplit(students(), "\n")[[1]]
    students_vector <- unique(students_vector) # remove duplicates from input
    N_students <- length(students_vector)
    
    if(N_students > N_spots) {
      
      sampled_students <- sample(students_vector, size = N_spots, replace = F)
      df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
      
    } else if (N_students < input$N_seats) { 
      
      df_results <- as.data.frame(matrix(students_vector, ncol = input$N_lectures, nrow = N_students))
      
    } else {
      
      N_visits_every_stud <- N_spots %/% N_students # number of times every student may come
      remaining_spots <- N_spots %% N_students # remaining number of spots to sample for
      
      if(remaining_spots > input$N_seats){
        # if the number of remaining spots bigger than N_seats, the pre-last lecture should be filled with a sample based on remaining students

        filled_students_last_lecture_big_remaining <- input$N_seats-(remaining_spots %% input$N_seats) 
        # number of seats in last filled lecture that ARE filled
        
        studnr_already_in_last_lecture_big_remaining <- students_vector[(N_students-filled_students_last_lecture_big_remaining+1):N_students]
        
        remaining_students_to_sample_from_big_remaining <- students_vector[!students_vector %in% studnr_already_in_last_lecture_big_remaining]
        remaining_spots_sample_big_remaining <- sample(remaining_students_to_sample_from_big_remaining, size = (input$N_seats-filled_students_last_lecture_big_remaining), replace = F) 
        # fill up pre-last lect with a sample of remaining_students_to_sample_from_big_remaining
        
        last_lecture_full_sample <- sample(students_vector, size = (input$N_seats), replace = F) # normal sample for last lecture
        
        sampled_students <- c(rep(students_vector, times= N_visits_every_stud), remaining_spots_sample_big_remaining, last_lecture_full_sample)
        # all together
        
        df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
        
      } else {
        filled_students_last_lecture <- input$N_seats-remaining_spots
        studnr_already_in_last_lecture <- students_vector[(N_students-filled_students_last_lecture+1):N_students]
        # collect studentnrs that are already placed in the last lecture
        
        remaining_students_to_sample_from <- students_vector[!students_vector %in% studnr_already_in_last_lecture]
        # remove the already placed students from the last lecture from students_vector, these are the remaining students for remaining_spots_sample
        
        remaining_spots_sample <- sample(remaining_students_to_sample_from, size = remaining_spots, replace = F)
        sampled_students <- c(rep(students_vector, times= N_visits_every_stud), remaining_spots_sample)
        
        df_results <- as.data.frame(matrix(sampled_students, ncol = input$N_lectures, nrow = input$N_seats))
      }
    }
    
    colnames(df_results) <- paste0("Lecture ", 1:input$N_lectures)
    df_results
    
  })
  
 # output$result <- renderTable({results()})
  
  output$download_table <- downloadHandler(
    filename = function(){paste("lecture-seats-sampler", Sys.Date(), ".xlsx", sep="")},
    content = function(file){writexl::write_xlsx(results(), file)},
    contentType = "text/csv")
  
  observeEvent(input$submit_button, {
    if (input$submit_button == FALSE){
      shinyjs::hide("download_table")}
    else {shinyjs::show("download_table")}
  })
```
